---
title: "P8105_hw6_yy3295"
author: "Ye Yuan"
date: "2023-12-01"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)

set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Question 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Question 2

Download the Central Park weather data.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Use 5000 bootstrap samples.

```{r}
boot_straps = 
  weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin, data = df)),
    results_1 = map(models, broom::tidy),
    results_2 = map(models, broom::glance)) |>
  select(-strap, -models)
```

For each bootstrap sample, produce estimates (r^2)

```{r}
r2 = 
  boot_straps |>
  select(-results_1) |>
  unnest(results_2) |>
  select(.id, r.squared)
```

Plot the distribution of your estimates (r^2)

```{r}
r2 |> 
  ggplot(aes(x = r.squared)) + geom_density()
```

From the plot we can tell that it is a left skewed distribution. 

Identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval (r^2)

```{r}
lower_quantile_r <- quantile(r2$r.squared, 0.025)

upper_quantile_r <- quantile(r2$r.squared, 0.975)

confidence_interval_r <- c(lower_quantile_r, upper_quantile_r)

confidence_interval_r
```
The 2.5% quantile is 0.886, and 97.5% quantile is 0.938.

The 95% confidence interval is (0.886, 0.938).

For each bootstrap sample, produce estimates (log(b1 * b2))

```{r}
logb = 
  boot_straps |>
  select(-results_2) |>
  unnest(results_1) |>
  select(.id, term, estimate) |>
  mutate(term = ifelse(term == "(Intercept)", "b1",
                       ifelse(term == "tmin", "b2", term))) |>
  pivot_wider(names_from = term,
              values_from = estimate) |>
  mutate(log_new = log(b1 * b2))
```

Plot the distribution of your estimates (log(b1 * b2))

```{r}
logb |> 
  ggplot(aes(x = log_new)) + geom_density()
```

From the plot we can tell that it is a normal distribution. 

Identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval (log(b1 * b2))

```{r}
lower_quantile_log <- quantile(logb$log_new, 0.025)

upper_quantile_log <- quantile(logb$log_new, 0.975)

confidence_interval_log <- c(lower_quantile_log, upper_quantile_log)

confidence_interval_log
```
The 2.5% quantile is 2.049, and 97.5% quantile is 2.133.

The 95% confidence interval is (2.049, 2.133).

### Question 3

Load and clean the data for regression analysis

```{r}
birthweight_df = 
  read_csv("data/birthweight.csv") |>
  janitor::clean_names() |>
  mutate(babysex = case_when(babysex == 1 ~ "male",
                         babysex == 2 ~ "female"),
         babysex = as.factor(babysex),
         frace = case_when(frace == 1 ~ "white",
                           frace == 2 ~ "black",
                           frace == 3 ~ "asian",
                           frace == 4 ~ "puerto_rican",
                           frace == 8 ~ "other",
                           frace == 9 ~ "unknown"),
         frace = as.factor(frace),
         malform = case_when(malform == 0 ~ "absent",
                             malform == 1 ~ "present"),
         malform = as.factor(malform),
         mrace = case_when(mrace == 1 ~ "white",
                           mrace == 2 ~ "black",
                           mrace == 3 ~ "asian",
                           mrace == 4 ~ "puerto_rican",
                           mrace == 8 ~ "other"),
         mrace = as.factor(mrace))

sum(is.na(birthweight_df))
```

